---
format: gfm
default-image-extension: ""
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
library(apa7)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# apa7

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/apa7)](https://CRAN.R-project.org/package=apa7)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![ggdiagram status badge](https://wjschne.r-universe.dev/apa7/badges/version)](https://wjschne.r-universe.dev/apa7)
<!-- badges: end -->

The goal of apa7 is to facilitate writing documents in APA Style (7th Edition).

## Installation

To install the published version from CRAN:

```{r}
#| eval: false
install.packages("apa7")
```


You can install the development version of apa7 with:

``` r
install.packages("apa7", repos = c('https://wjschne.r-universe.dev'))
```

## Tables

The package provides functions to create APA-style tables, including correlation matrices and regression tables. The tables can be formatted using the `flextable` package.

```{r}
#| label: showcor
#| message: false
#| warning: false
#| fig-alt: A correlation matrix in APA style.
library(apa7)
# Correlation matrix 
apa_cor(trees, star_significant = TRUE)

```


```{r}
#| label: showlm
#| message: false
#| warning: false
#| fig-alt: A regression table in APA style.
# Make regression model, format parameters, and display flextable
lm(Volume ~ Girth + Height, data = trees) |> 
  apa_parameters() |> 
  apa_flextable()
```

```{r}
#| label: showcontingence
#| message: false
#| warning: false
#| fig-alt: A contigency table with a chi-square test of independence e in APA style.
# Contingency table with chi-square test of independence
d <- mtcars[, c("am", "gear")]
colnames(d) <- c("Transmission", "Gears")
d$Transmission <- factor(d$Transmission, 
                         levels = c(0, 1), 
                         labels = c("Automatic", "Manual"))
apa_chisq(d)
```


## Formatting functions

The package provides functions to format p-values, numbers, and other statistical results according to APA Style.

```{r}
#| label: formatters
# Format p-values
apa_p(c(0.0007, 0.001, 0.0081, 0.024, 0.454))
```

