---
title: "Correlation Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(apa7)
library(dplyr)
```

Some tables, like correlation matrices, can be standardized and automated. Others are uniquely tied to a specific purpose, and no function could anticipate their structure. The apa7 package has a few functions to automate what can be automated and a styling function to make the final styling of a complex table a little easier.

By default, the output of `apa_cor` [flextable](https://davidgohel.github.io/flextable/), with significant correlations bolded.

```{r}
trees %>% 
  apa_cor()
```

If you prefer to work with [gt tables](https://gt.rstudio.com/), set `output` function to `gt`.

```{r}
trees %>% 
  apa_cor(output = "gt") 
```

If you want to use some other table package, you can output a tibble instead, and then process the table however you want.

```{r}
options(knitr.kable.NA = '')
trees %>% 
  apa_cor(output = "tibble") %>% 
  knitr::kable(align = "lrrrrr")
```

# Styling options

Your styling options include:

* `note` A custom markdown formatted note. Overrides automatic note about bolded correlations being significant. Default: `NULL`
* `p_value` Significance value. Default: `.05`
* `digits` Number of digits for rounding. Default: `2`
* `bold_significant` Whether to bold significant correlations. Default:  `TRUE`
* `significance_note` Whether note about bolding correlations appears. Overridden by a custom note. Also, the automatic note will not appear if no correlations are significant. Default:  `TRUE`
* `output` Options are `flextable`, `gt` or `tibble`. Default: `flextable`
* `family` Font family (typeface). Default: `Times New Roman`
* `font_size` Font size (in points). Default: `12`
* `text_color` Text color. Default: `black`
* `border_color` Table border color. Default: `black`
* `border_width` Table border width. Default: `0.5`
* `line_spacing` Text line spacing. Default: `2`


For example, 

```{r}
# List of descriptive functions
my_summaries <- list(M = mean,
                     SD = sd,
                     Skew = psych::skew,
                     `Kurt.` = psych::kurtosi,
                     n = function(x) sum(!is.na(x)))
trees %>% 
  apa_cor(note = "*M* = Mean, *SD* = Standard Deviation, *Skew* =  Skewness, *Kurt.* =  Kurtosis",
          p_value = .001, 
          digits = 2,
          bold_significant = FALSE, 
          significance_note = FALSE,
          summary_functions = my_summaries,
          output = "gt", 
          family = "Arial", 
          font_size = 10, 
          text_color = "darkred",
          border_color = "royalblue4", 
          border_width = 1, 
          line_spacing = 2)
```

To remove summary statistics entirely, set `summary_functions` to `NULL`.

```{r}
apa_cor(trees, 
        summary_functions = NULL, 
        bold_significant = FALSE)
```

# Customization

The styling options for `apa_cor` are minimal, but flexable and gt tables allow for all kinds of customization. Just place whatever code from flextable or gt you want after `apa_cor`. For example, gt tables usually span 100% of the width in a space. Here I set the width to 50% and align the table to the left of the page. I place a footnote in a specific cell in the table body.

```{r}
apa_cor(trees, output = "gt", significance_note = F) %>% 
  gt::tab_options(table.width = gt::pct(50), 
                  table.align = "left") %>% 
  gt::tab_footnote("This is a footnote.", 
                   locations = gt::cells_body(columns = "Variable", 
                                              rows = 2))
```

Here I do the same with flextable:

```{r}
apa_cor(trees, significance_note = F) %>%
  flextable::set_table_properties(align = "left") %>%
  flextable::width(j = c("Variable", "M", "SD", "1", "2", "3"), width = c(1, .45, .45, .45, .45, .45)) %>% 
  flextable::footnote(
    i = 2,
    j = "Variable",
    value = flextable::as_paragraph("This is a footnote."),
    part = "body"
  ) 

```

